Concentration of Maxima and Fundamental Limits in High-Dimensional Testing and Inference

### 2021-02-04 5.URS/5.proofs-UDD-URS.tex

\begin{proof}[Proof of Lemma \ref{lemma:AGG-maxima-upper-tails}]
Recall by \eqref{eq:AGG-quantiles} that 
\begin{equation*}
    u_q\sim\left(\nu\log{q}\right)^{1/\nu}, \quad q\to\infty,
\end{equation*}
so that
\begin{equation} %\label{eq:choice-of-c_p}
c_p  = \frac{u_{p\log{p}}}{u_p} -1 = \left(\frac{\log{p}+\log{\log{p}}}{\log{p}}\right)^{1/\nu}(1+o(1)) - 1 \rightarrow 0 \quad \mbox{as } p\to\infty.
\end{equation} 
By the union bound, we have
\begin{align}
    \P\left[\frac{M_p}{u_p} > 1+c_p\right] 
        &\le \sum_{j=1}^p \P\left[\frac{\epsilon_p(j)}{u_p} > 1+c_p\right] 
        = p \overline{F}\left(u_{p\log{p}}\right) \label{eq:AGG-maxima-upper-tails-proof-1} \\
        &= p \overline{F}\left(F^{\leftarrow}\left(1-\frac{1}{p\log{p}}\right)\right) \le \frac{1}{\log{p}} \rightarrow 0. \nonumber
\end{align}
where the last inequality follows from the fact that $F\left(F^{\leftarrow}(u)\right)\ge u$ for all $u\in[0,1]$.
\end{proof}

@@

\begin{lemma} \label{lemma:AGG-max-uniform-integrability}
Let $M_p$ and $c_p$ be as in Lemma \ref{lemma:AGG-maxima-upper-tails}, and denote 
$$
\xi_p := \frac{M_p}{(1+c_p)u_p}.
$$
Then there exist $p_0, t_0 > 0$, and an absolute constant $C>0$ such that
\begin{equation} \label{eq:lemma-AGG-uniform-integrability}
    \P\left[\xi_p > t \right]\le \exp{\{-Ct^\nu\}}, \quad \text{for all} \quad p>p_0,\ t>t_0.
\end{equation}
In particular, the set of random variables $\{\left(\xi_p\right)_+,p\in\N\}$ is uniformly integrable.
\end{lemma}
\begin{proof}[Proof of Lemma \ref{lemma:AGG-max-uniform-integrability}]
Recalling that $(1+c_p)u_p = u_{p\log{p}}$, and by applying the union bound as in \eqref{eq:AGG-maxima-upper-tails-proof-1}, we have
\begin{align}
    \log \P\left[\xi_p > t\right] 
        &\le \log p + \log{\overline{F}\left(u_{p\log{p}}t\right)} \nonumber \\
        &\le \log p - \frac{1}{\nu}\left(u_{p\log{p}}t\right)^\nu(1-\delta). \label{eq:lemma-AGG-uniform-integrability-proof-1}
\end{align}
for $t > t_0(\delta)>0$, where $\delta\in(0,1)$ is an arbitrarily small number fixed in advance. 
This follows from the assumption that $F\in\text{AGG}(\nu)$ and Definition \ref{def:AGG} of the AGG distribution.
Using in \eqref{eq:lemma-AGG-uniform-integrability-proof-1} the explicit expressions for the quantiles in \eqref{eq:AGG-quantiles}, we obtain
\begin{equation} \label{eq:lemma-AGG-uniform-integrability-proof-2}
    \log \P\left[\xi_p > t\right] \le \log p - \underbrace{\left(1+o(1)\right)(1-\delta)t^\nu}_{\text{greater than }1\text{ for large }t}\log{p} - t^\nu\underbrace{\log{\log{p}}\left(1+o(1)\right)(1-\delta)}_{\text{greater than }C\text{ for large }p}.
\end{equation}
For large $t$, we have $\left(1+o(1)\right)(1-\delta)t^\nu > 1$, so that the sum of the first two terms on the right-hand side of \eqref{eq:lemma-AGG-uniform-integrability-proof-2} is negative.
Also, for $p$ larger than some constant $p_0(\delta)$, we have $\log{\log{p}}\left(1+o(1)\right)(1-\delta) > C$ for some constant $C$ that does not depend on $p$.
Therefore \eqref{eq:lemma-AGG-uniform-integrability} holds for $t>t_0(\delta)$ and $p>p_0(\delta)$, and the proof is complete.
\end{proof}

@@

\begin{proof}[Proof of Corollary \ref{cor:AGG-max-upper-bound-expectation}]
Since $c_p\ge0$ is a sequence converging to 0, we have $c_p < 1$ for $p \ge p_0$. Hence for any $t>0$, we have
\begin{align}
    \P\left[\left(\frac{M_p}{u_p} - (1+c_p)\right)_+ > t\right] 
    &= \P\left[(1+c_p)\left(\xi_p-1\right)_+ > t\right] \nonumber \\
    &\le \P\left[\left(\xi_p-1\right)_+ > t/2\right] 
    \le \P\left[\xi_p > t/2 \right]. \label{eq:AGG-max-bound-upper-expectation-proof}
\end{align}
By Lemma \ref{lemma:AGG-max-uniform-integrability}, $\{\left(\xi_p\right)_+\}$ is u.i., therefore by Relation \eqref{eq:AGG-max-bound-upper-expectation-proof}, $\{\left(M_p/u_p - (1+c_p)\right)_+,\;p\in\N\}$ is u.i. as well.
Since by Lemma \ref{lemma:AGG-maxima-upper-tails}, $\left(M_p/u_p - (1+c_p)\right)_+\to 0$ in probability, Relation \eqref{eq:AGG-max-bound-upper-expectation} follows from the established uniform integrability \citep[see, e.g., Theorem 6.6.1 in][]{resnick:1999book}.
\end{proof}

@@

\begin{proof}[Proof of Lemma \ref{lemma:Gaussian-maxima-lower-expectation}]
We start with the identity
$$
\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right] = \E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_+\right] - \E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_-\right].
$$
By re-arranging terms and taking limsup/liminf, we obtain
\begin{align} \label{e:lim-sup-vanishes}
    0 \le &\limsup_{p\to\infty}\E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_-\right]\nonumber \\
        \le & \limsup_{p\to\infty}\E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_+\right] - \liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right]\\
        = & - \liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right],
        \label{e:lim-inf-bound}
\end{align}
where the last equality follows from the fact that the lim-sup in \eqref{e:lim-sup-vanishes} vanishes by Corollary \ref{cor:AGG-max-upper-bound-expectation}.
On the other hand, since $c_q\to 0$, we have
$$
\liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right] 
= \liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-1\right] \ge 0,
$$
where the last inequality follows from Proposition \ref{prop:Gaussian-maxima-expectation-lower-bound}.  This shows that 
the right-hand side of \eqref{e:lim-inf-bound} is non-positive and hence
\eqref{eq:Gaussian-maxima-lower-expectation} holds. 
\end{proof}

>>>>

\begin{proof}[Lemma \ref{lemma:AGG-maxima-upper-tails}]
Recall by \eqref{eq:AGG-quantiles} that 
\begin{equation*}
    u_q\sim\left(\nu\log{q}\right)^{1/\nu}, \quad q\to\infty,
\end{equation*}
so that
\begin{equation} %\label{eq:choice-of-c_p}
c_p  = \frac{u_{p\log{p}}}{u_p} -1 = \left(\frac{\log{p}+\log{\log{p}}}{\log{p}}\right)^{1/\nu}(1+o(1)) - 1 \rightarrow 0 \quad \mbox{as } p\to\infty.
\end{equation} 
By the union bound, we have
\begin{align}
    \P\left[\frac{M_p}{u_p} > 1+c_p\right] 
        &\le \sum_{j=1}^p \P\left[\frac{\epsilon_p(j)}{u_p} > 1+c_p\right] 
        = p \overline{F}\left(u_{p\log{p}}\right) \label{eq:AGG-maxima-upper-tails-proof-1} \\
        &= p \overline{F}\left(F^{\leftarrow}\left(1-\frac{1}{p\log{p}}\right)\right) \le \frac{1}{\log{p}} \rightarrow 0. \nonumber
\end{align}
where the last inequality follows from the fact that $F\left(F^{\leftarrow}(u)\right)\ge u$ for all $u\in[0,1]$.
\end{proof}

@@

\begin{lemma} \label{lemma:AGG-max-uniform-integrability}
Let $M_p$ and $c_p$ be as in Lemma \ref{lemma:AGG-maxima-upper-tails}, and denote 
$$
\xi_p := \frac{M_p}{(1+c_p)u_p}.
$$
Then there exist $p_0, t_0 > 0$, and an absolute constant $C>0$ such that
\begin{equation} \label{eq:lemma-AGG-uniform-integrability}
    \P\left[\xi_p > t \right]\le \exp{\{-Ct^\nu\}}, \quad \text{for all} \quad p>p_0,\ t>t_0.
\end{equation}
In particular, the set of random variables $\{\left(\xi_p\right)_+,p\in\N\}$ is uniformly integrable.
\end{lemma}
\begin{proof}[Lemma \ref{lemma:AGG-max-uniform-integrability}]
Recalling that $(1+c_p)u_p = u_{p\log{p}}$, and by applying the union bound as in \eqref{eq:AGG-maxima-upper-tails-proof-1}, we have
\begin{align}
    \log \P\left[\xi_p > t\right] 
        &\le \log p + \log{\overline{F}\left(u_{p\log{p}}t\right)} \nonumber \\
        &\le \log p - \frac{1}{\nu}\left(u_{p\log{p}}t\right)^\nu(1-\delta). \label{eq:lemma-AGG-uniform-integrability-proof-1}
\end{align}
for $t > t_0(\delta)>0$, where $\delta\in(0,1)$ is an arbitrarily small number fixed in advance. 
This follows from the assumption that $F\in\text{AGG}(\nu)$ and Definition \ref{def:AGG} of the AGG distribution.
Using in \eqref{eq:lemma-AGG-uniform-integrability-proof-1} the explicit expressions for the quantiles in \eqref{eq:AGG-quantiles}, we obtain
\begin{equation} \label{eq:lemma-AGG-uniform-integrability-proof-2}
    \log \P\left[\xi_p > t\right] \le \log p - \underbrace{\left(1+o(1)\right)(1-\delta)t^\nu}_{\text{greater than }1\text{ for large }t}\log{p} - t^\nu\underbrace{\log{\log{p}}\left(1+o(1)\right)(1-\delta)}_{\text{greater than }C\text{ for large }p}.
\end{equation}
For large $t$, we have $\left(1+o(1)\right)(1-\delta)t^\nu > 1$, so that the sum of the first two terms on the right-hand side of \eqref{eq:lemma-AGG-uniform-integrability-proof-2} is negative.
Also, for $p$ larger than some constant $p_0(\delta)$, we have $\log{\log{p}}\left(1+o(1)\right)(1-\delta) > C$ for some constant $C$ that does not depend on $p$.
Therefore \eqref{eq:lemma-AGG-uniform-integrability} holds for $t>t_0(\delta)$ and $p>p_0(\delta)$, and the proof is complete.
\end{proof}

@@

\begin{proof}[Corollary \ref{cor:AGG-max-upper-bound-expectation}]
Since $c_p\ge0$ is a sequence converging to 0, we have $c_p < 1$ for $p \ge p_0$. Hence for any $t>0$, we have
\begin{align}
    \P\left[\left(\frac{M_p}{u_p} - (1+c_p)\right)_+ > t\right] 
    &= \P\left[(1+c_p)\left(\xi_p-1\right)_+ > t\right] \nonumber \\
    &\le \P\left[\left(\xi_p-1\right)_+ > t/2\right] 
    \le \P\left[\xi_p > t/2 \right]. \label{eq:AGG-max-bound-upper-expectation-proof}
\end{align}
By Lemma \ref{lemma:AGG-max-uniform-integrability}, $\{\left(\xi_p\right)_+\}$ is u.i., therefore by Relation \eqref{eq:AGG-max-bound-upper-expectation-proof}, $\{\left(M_p/u_p - (1+c_p)\right)_+,\;p\in\N\}$ is u.i. as well.
Since by Lemma \ref{lemma:AGG-maxima-upper-tails}, $\left(M_p/u_p - (1+c_p)\right)_+\to 0$ in probability, Relation \eqref{eq:AGG-max-bound-upper-expectation} follows from the established uniform integrability \citep[see, e.g., Theorem 6.6.1 in][]{resnick:1999book}.
\end{proof}

@@

\begin{proof}[Lemma \ref{lemma:Gaussian-maxima-lower-expectation}]
We start with the identity
$$
\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right] = \E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_+\right] - \E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_-\right].
$$
By re-arranging terms and taking limsup/liminf, we obtain
\begin{align} \label{e:lim-sup-vanishes}
    0 \le &\limsup_{p\to\infty}\E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_-\right]\nonumber \\
        \le & \limsup_{p\to\infty}\E\left[\left(\frac{M_{S_p}}{u_q}-(1+c_q)\right)_+\right] - \liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right]\\
        = & - \liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right],
        \label{e:lim-inf-bound}
\end{align}
where the last equality follows from the fact that the lim-sup in \eqref{e:lim-sup-vanishes} vanishes by Corollary \ref{cor:AGG-max-upper-bound-expectation}.
On the other hand, since $c_q\to 0$, we have
$$
\liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-(1+c_q)\right] 
= \liminf_{p\to\infty}\E\left[\frac{M_{S_p}}{u_q}-1\right] \ge 0,
$$
where the last inequality follows from Proposition \ref{prop:Gaussian-maxima-expectation-lower-bound}.  This shows that 
the right-hand side of \eqref{e:lim-inf-bound} is non-positive and hence
\eqref{eq:Gaussian-maxima-lower-expectation} holds. 
\end{proof}

### 2021-02-04 6.GWAS/6.chi-square-distributions.tex

\begin{proof}[Proof of Lemma \ref{lemma:rapid-variation-chisq}]
When $\nu=1$, the chi-square distribution reduces to a squared Normal, and \eqref{eq:rapid-variation-chisq} follows from the rapid variation of the standard Normal distribution.
For $\nu\ge2$, we recall the following bound on tail probabilities (see, e.g., \citep{inglot2010inequalities}),
$$
\frac{1}{2}\mathcal{E}_\nu(x) \le \P[\chi_\nu^2(0)>x] \le \frac{x}{(x-\nu+2)\sqrt{\pi}} \mathcal{E}_\nu(x), \quad \nu\ge2,\;x>\nu-2,
$$
where $\mathcal{E}_\nu(x) = \exp\left\{-\frac{1}{2}[(x-\nu-(\nu-2)\log(x/\nu) + \log\nu]\right\}$.
Therefore, we have 
$$
\frac{(x-\nu+2)\sqrt{\pi}}{2x}\frac{\mathcal{E}_\nu(tx)}{\mathcal{E}_\nu(x)} 
\le \frac{\P[\chi_\nu^2(0)>tx]}{\P[\chi_\nu^2(0)>x]}
\le \frac{2tx}{(tx-\nu+2)\sqrt{\pi}}\frac{\mathcal{E}_\nu(tx)}{\mathcal{E}_\nu(x)},
$$
where ${\mathcal{E}_\nu(tx)}/{\mathcal{E}_\nu(x)} = \exp\{-\frac{1}{2}[(t-1)x-(\nu-2)\log{t}]\}$ converges to $0$ or $\infty$ depending on whether $t>1$ or $0<t<1$.
The case where $t=1$ is trivial.
\end{proof}

@@

\begin{proof}[Proof of Lemma \ref{lemma:stochastic-monotonicity}]
Recall that non-central chi-square distributions can be written as sums of $\nu-1$ standard normal random variables and a non-central normal random variable with mean $\sqrt{\lambda}$ and variance 1,
\begin{equation*}
    \chi_\nu^2(\lambda) 
    \stackrel{\mathrm{d}}{=} Z_1^2 + \ldots + Z_{\nu-1}^2 + (Z_\nu + \sqrt{\lambda})^2.
\end{equation*}
Therefore, it suffices to show that $\P[(Z+\sqrt{\lambda})^2 \le t]$ is non-increasing in $\lambda$ for any $t\ge0$, where $Z$ is a standard normal random variable.
We rewrite this expression in terms of standard normal probability function $\Phi$,
\begin{align}
    \P[(Z+\sqrt{\lambda})^2 \le t] 
    &= \P[-\sqrt{\lambda} - \sqrt{t} \le Z \le -\sqrt{\lambda} + \sqrt{t}] \nonumber \\
    &= \Phi(-\sqrt{\lambda} + \sqrt{t}) - \Phi(-\sqrt{\lambda} - \sqrt{t}). \label{eq:stochastic-monotonicity-proof-1}
\end{align}
The derivative of the last expression (with respect to $\lambda$) is 
\begin{equation} \label{eq:stochastic-monotonicity-proof-2}
    \frac{1}{2\sqrt{\lambda}} \left(\phi(\sqrt{\lambda} + \sqrt{t}) - \phi(\sqrt{\lambda} - \sqrt{t})\right) 
    = \frac{1}{2\sqrt{\lambda}} \left(\phi(\sqrt{\lambda} + \sqrt{t}) - \phi(\sqrt{t} - \sqrt{\lambda})\right),
\end{equation}
where $\phi$ is the density of the standard normal distribution.
Notice that we have used the symmetry of $\phi$ around 0 in the last expression.

@@

\begin{proof}[Proof of Lemma \ref{lemma:chisq-quantiles}]
The case where $\nu=1$ follows from the well-known formula for Normal quantiles (see, e.g., Proposition 1.1 in \cite{gao2018fundamental})
$$
F^\leftarrow(1 - y) = \Phi^\leftarrow(1-y/2)\sim\sqrt{2\log{(2/y)}}\sim\sqrt{2\log{(1/y)}}.
$$
The case where $\nu\ge2$ follows from the following estimates of high quantiles of chi-square distributions (see, e.g., \citep{inglot2010inequalities}),
$$
    \nu +  2\log(1/y) -5/2 \le u(y) \le \nu +  2\log(1/y) + 2\sqrt{\nu\log(1/y)}, \quad \text{for all }y\le0.17,
$$
where both the lower and upper bound are asymptotic to $2\log(1/y)$.
\end{proof}

>>>>

\begin{proof}[Lemma \ref{lemma:rapid-variation-chisq}]
When $\nu=1$, the chi-square distribution reduces to a squared Normal, and \eqref{eq:rapid-variation-chisq} follows from the rapid variation of the standard Normal distribution.
For $\nu\ge2$, we recall the following bound on tail probabilities (see, e.g., \citep{inglot2010inequalities}),
$$
\frac{1}{2}\mathcal{E}_\nu(x) \le \P[\chi_\nu^2(0)>x] \le \frac{x}{(x-\nu+2)\sqrt{\pi}} \mathcal{E}_\nu(x), \quad \nu\ge2,\;x>\nu-2,
$$
where $\mathcal{E}_\nu(x) = \exp\left\{-\frac{1}{2}[(x-\nu-(\nu-2)\log(x/\nu) + \log\nu]\right\}$.
Therefore, we have 
$$
\frac{(x-\nu+2)\sqrt{\pi}}{2x}\frac{\mathcal{E}_\nu(tx)}{\mathcal{E}_\nu(x)} 
\le \frac{\P[\chi_\nu^2(0)>tx]}{\P[\chi_\nu^2(0)>x]}
\le \frac{2tx}{(tx-\nu+2)\sqrt{\pi}}\frac{\mathcal{E}_\nu(tx)}{\mathcal{E}_\nu(x)},
$$
where ${\mathcal{E}_\nu(tx)}/{\mathcal{E}_\nu(x)} = \exp\{-\frac{1}{2}[(t-1)x-(\nu-2)\log{t}]\}$ converges to $0$ or $\infty$ depending on whether $t>1$ or $0<t<1$.
The case where $t=1$ is trivial.
\end{proof}

@@

\begin{proof}[Lemma \ref{lemma:stochastic-monotonicity}]
Recall that non-central chi-square distributions can be written as sums of $\nu-1$ standard normal random variables and a non-central normal random variable with mean $\sqrt{\lambda}$ and variance 1,
\begin{equation*}
    \chi_\nu^2(\lambda) 
    \stackrel{\mathrm{d}}{=} Z_1^2 + \ldots + Z_{\nu-1}^2 + (Z_\nu + \sqrt{\lambda})^2.
\end{equation*}
Therefore, it suffices to show that $\P[(Z+\sqrt{\lambda})^2 \le t]$ is non-increasing in $\lambda$ for any $t\ge0$, where $Z$ is a standard normal random variable.
We rewrite this expression in terms of standard normal probability function $\Phi$,
\begin{align}
    \P[(Z+\sqrt{\lambda})^2 \le t] 
    &= \P[-\sqrt{\lambda} - \sqrt{t} \le Z \le -\sqrt{\lambda} + \sqrt{t}] \nonumber \\
    &= \Phi(-\sqrt{\lambda} + \sqrt{t}) - \Phi(-\sqrt{\lambda} - \sqrt{t}). \label{eq:stochastic-monotonicity-proof-1}
\end{align}
The derivative of the last expression (with respect to $\lambda$) is 
\begin{equation} \label{eq:stochastic-monotonicity-proof-2}
    \frac{1}{2\sqrt{\lambda}} \left(\phi(\sqrt{\lambda} + \sqrt{t}) - \phi(\sqrt{\lambda} - \sqrt{t})\right) 
    = \frac{1}{2\sqrt{\lambda}} \left(\phi(\sqrt{\lambda} + \sqrt{t}) - \phi(\sqrt{t} - \sqrt{\lambda})\right),
\end{equation}
where $\phi$ is the density of the standard normal distribution.
Notice that we have used the symmetry of $\phi$ around 0 in the last expression.

@@

\begin{proof}[Lemma \ref{lemma:chisq-quantiles}]
The case where $\nu=1$ follows from the well-known formula for Normal quantiles (see, e.g., Proposition 1.1 in \cite{gao2018fundamental})
$$
F^\leftarrow(1 - y) = \Phi^\leftarrow(1-y/2)\sim\sqrt{2\log{(2/y)}}\sim\sqrt{2\log{(1/y)}}.
$$
The case where $\nu\ge2$ follows from the following estimates of high quantiles of chi-square distributions (see, e.g., \citep{inglot2010inequalities}),
$$
    \nu +  2\log(1/y) -5/2 \le u(y) \le \nu +  2\log(1/y) + 2\sqrt{\nu\log(1/y)}, \quad \text{for all }y\le0.17,
$$
where both the lower and upper bound are asymptotic to $2\log(1/y)$.
\end{proof}

### 2021-02-04 6.GWAS/6.odds-and-power.tex

Proposition \ref{prop:signal-size-odds-ratio} is derived in Section \ref{subsec:proof-signal-size-odds-ratio}.

@@

>>>>

\begin{proof}
	We parametrize the 2-by-2 multinomial distribution with the parameter $\delta$, 
	\begin{equation} \label{eq:reparametrize-2-by-2-table-1}
	\mu_{11} = \phi_1\theta_1+\delta,\quad 
	\mu_{12} = \phi_1\theta_2-\delta,\quad 
	\mu_{21} = \phi_2\theta_1-\delta,\quad 
	\mu_{22} = \phi_2\theta_2+\delta.
	\end{equation}
	By relabelling of categories, we may assume $0<\theta_1,\phi_1\le1/2$ without loss of generality.
	Note that $\delta$ must lie within the range $[\delta_\mathrm{min}, \delta_\mathrm{max}]$, where
	$$
	\delta_\mathrm{min} := \max\{-\phi_1\theta_1, -\phi_2\theta_2, \phi_1\theta_2-1, \phi_2\theta_1-1\} 
	= -\phi_1\theta_1,
	$$
	and
	$$
	\delta_\mathrm{max} := \min\{1-\phi_1\theta_1, 1-\phi_2\theta_2, \phi_1\theta_2, \phi_2\theta_1\}
	= \min\{\phi_1\theta_2, \phi_2\theta_1\},
	$$
	in order for $\mu_{ij}\ge0$ for all $i,j\in \{1,2\}$.
	Under this parametrization, Relation \eqref{eq:odds-ratio} then becomes
	\begin{equation} \label{eq:odds-ratio-delta}
	\text{R} = \frac{\mu_{11}\mu_{22}}{\mu_{12}\mu_{21}}
	= \frac{\phi_1\theta_1\phi_2\theta_2 + \delta(\phi_1\theta_1+\phi_2\theta_2)+\delta^2}{\phi_1\theta_1\phi_2\theta_2 - \delta(\phi_1\theta_2+\phi_2\theta_1)+\delta^2},
	\end{equation}
	which is one-to-one and increasing in $\delta$ on $(\delta_\mathrm{min}, \delta_\mathrm{max})$.
	Equation \eqref{eq:signal-size-chisq} becomes
	\begin{equation} \label{eq:signal-size-chisq-delta}
	w^2 = \sum_{i=1}^2 \sum_{j=1}^2 \frac{(\mu_{ij} - \phi_i\theta_j)^2}{\phi_i\theta_j}
	= \delta^2\sum_i\sum_j \frac{1}{\phi_i\theta_j}
	= \frac{\delta^2}{\phi_1\theta_1\phi_2\theta_2},
	\end{equation}
	Solving for $\delta$ in \eqref{eq:odds-ratio-delta}, and plugging into the expression for signal size \eqref{eq:signal-size-chisq-delta} yields Relation \eqref{eq:signal-size-odds-ratio}.

	The other three cases ($1/2\le\theta_1,\phi_1\le1$; $0<\theta_1\le1/2\le\phi_1\le1$; and $0\le\phi_1\le1/2\le\theta_1\le1$) may be obtained similarly, or by appealing to the symmetry of the problem.
\end{proof}

@@

\begin{proof}
	As in the proof of Proposition \ref{prop:signal-size-odds-ratio}, we examine the case where $0<\theta_1,\phi_1\le1/2$, and leave the other three cases an exercise.
	Take the first derivative of the expression for $w^2$ in equation \eqref{eq:signal-size-chisq-delta} with respect to $\delta$, it is evident that $w^2(\delta)$ is decreasing on $[\delta_\mathrm{min},0)$, increasing on $(0,\delta_\mathrm{max}]$, with limits
	$$
	\lim_{d\to \delta_\mathrm{min}} w^2(\delta) = \frac{\phi_1\theta_1}{\phi_2\theta_2},
	\quad
	\text{and}
	\quad
	\lim_{d\to \delta_\mathrm{max}} w^2(\delta) = \min\left\{\frac{\phi_1\theta_2}{\phi_2\theta_1}, \frac{\phi_2\theta_1}{\phi_1\theta_2}\right\}.
	$$
\end{proof}

### 2021-02-04 preamble.tex

{\bf The contributions of this work.} The fundamental limits of the classic detection problem hinge 
of the analysis of the discrepancy between the {\em null} and {\em alternative} hypotheses e.g., via Hellinger distance.  Thus, perhaps 
for  technical reasons, much of the analysis in the existing literature has been done under the assumption that the additive errors are 
independent and/or Gaussian, or using loss functions unaffected by the dependence such as the Hamming loss.  
In this work, we demonstrate that the support recovery problems, especially {\em exact support recovery}, are best understood from 
the novel perspective of the {\em concentration of maxima} phenomenon in extreme value theory.  It turns out that under a very broad 
range of light-tailed error distributions and under a {\em very} broad range of error dependence structures, the maxima of the errors, 
when rescaled (but not centered!) converge in probability to a positive constant. This concentration property leads to a complete solution 
of the exact  support recovery problem for the broad family of thresholding procedures.  Most if not all existing support estimation 
procedures are types thresholding procedures (see Section \ref{sec:statistical-procedures}).   
That is, the signal support estimate comprises of all components exceeding a suitable (potentially data-dependent)  threshold.  We show, by exploiting concentration of maxima, that thresholding procedures obey a phase-transition, where if the signal is above a 
certain boundary, asymptotically exact recovery is possible while below the boundary all thresholding procedures fail, as $p\to\infty$.  
Remarkably, light-tailed maxima concentrate under very broad and strong dependence.  This is exemplified by our characterization of
the concentration of maxima phenomenon for Gaussian triangular arrays.  For example, in the special case of stationary Gaussian time
series, vanishing auto-covariance is necessary and sufficient for the maxima to concentrate in the same way as independent standard 
normal random variables.   \stilian{This is in stark contrast with the behavior of sums, commonly studied under short- and long-range dependence conditions \citep[see e.g.][]{dedecker:doukhan:etal:2007,pipiras:taqqu:2017}.  Simply put, the notion of weak dependence that entails that the maxima of dependent variables concentrate at the
same rate as in the case of independence is fundamentally weaker than the conventional mixing conditions widely used in the study of sums.}{\fbox{added}}

\stilian{Our}{\xout{These}} probabilistic contributions may be of independent interest and extend classic work of \cite{berman1964limit}. 
Concentration of maxima is a type of super-concentration phenomenon studied also in \cite{chatterjee2014superconcentration} 
and \cite{tanguy:2015}. The robustness of the concentration of maxima phenomenon to dependence can perhaps explain the universality 
of phase transitions in support recovery problems.

>>>>

{\bf The contributions of this work.} The fundamental limits of the classic detection problem hinge 
of the analysis of the discrepancy between the {\em null} and {\em alternative} hypotheses e.g., via Hellinger distance.  Thus, perhaps 
for  technical reasons, much of the analysis in the existing literature has been done under the assumption that the additive errors are 
independent and/or Gaussian, or using loss functions unaffected by the dependence such as the Hamming loss.  
In this work, we demonstrate that the support recovery problems, especially {\em exact support recovery}, are best understood from 
the novel perspective of the {\em concentration of maxima} phenomenon in extreme value theory.  It turns out that under a very broad 
range of light-tailed error distributions and under a {\em very} broad range of error dependence structures, the maxima of the errors, 
when rescaled (but not centered!) converge in probability to a positive constant. This concentration property leads to a complete solution 
of the exact  support recovery problem for the broad family of thresholding procedures.  Most if not all existing support estimation 
procedures are types thresholding procedures (see Section \ref{sec:statistical-procedures}).   
That is, the signal support estimate comprises of all components exceeding a suitable (potentially data-dependent)  threshold.  We show, by exploiting concentration of maxima, that thresholding procedures obey a phase-transition, where if the signal is above a 
certain boundary, asymptotically exact recovery is possible while below the boundary all thresholding procedures fail, as $p\to\infty$.  
Remarkably, light-tailed maxima concentrate under very broad and strong dependence.  This is exemplified by our characterization of
the concentration of maxima phenomenon for Gaussian triangular arrays.  For example, in the special case of stationary Gaussian time
series, vanishing auto-covariance is necessary and sufficient for the maxima to concentrate in the same way as independent standard 
normal random variables.  This is in stark contrast with the behavior of sums, commonly studied under short- and long-range dependence conditions \citep[see e.g.][]{dedecker:doukhan:etal:2007,pipiras:taqqu:2017}.  Simply put, the notion of weak dependence that entails that the maxima of dependent variables concentrate at the
same rate as in the case of independence is fundamentally weaker than the conventional mixing conditions widely used in the study of sums.

Our probabilistic contributions may be of independent interest and extend classic work of \cite{berman1964limit}. 
Concentration of maxima is a type of super-concentration phenomenon studied also in \cite{chatterjee2014superconcentration} 
and \cite{tanguy:2015}. The robustness of the concentration of maxima phenomenon to dependence can perhaps explain the universality 
of phase transitions in support recovery problems.
